\LARGE \bf Enriching Robot's Actions with Affective Movements

### 2017-03-26 RO-MAN-2017-1/Parts/CaseStudy.tex
two ro-man articles compiled, ro-man2 still requires more work. let's see what the reviewers say about both articles

 The case study presented in this paper was held at an exhibition with two main objectives:
\begin{enumerate}
	\item Cross-validate the findings obtained from a previous experiment~\cite{Angel2017-2}, which was done to study the attribution of different linear and angular velocities, oscillation angle, direction, and orientation of the platform to express: \textit{Anger}, \textit{Happiness}, \textit{Sadness} and \textit{Fear}. A total of 196 value combinations were designed, 20 of which, randomly selected, were presented to each participant. Subjects had to select for each presentation the most representative term describing it among the four emotions and two mental states (i.e., \textit{Excitement} and \textit{Tenderness}), added as confounding terms.
	\item Verify whether participants would prefer scenes where the robot moves expressing emotions, or rather moves on the same trajectory without any emotion expression. To control scene variability two cameras and eight AR tags were used together with a Kalman filter to improve robot localization on the stage. AR tags were detected using the ROS package ar\_track\_alvar~\cite{artag2015}. The distribution of the cameras and the tags is reported in Figure~\ref{fig:setup_fourth}.

@@

\end{enumerate}

@@

This case study was done at an exhibition during a period of two days. People coming to the stand were asked to volunteerto participate in this study, which was divided in to parts. In the first one, each subject was exposed to two emotion expressions. %TODO Was it so? I have turned the sentence... 
The two emotions were generated to guarantee a random uniform presentation. In the second part, participants were explained that a small scene was going to be presented twice, and they had to select the one they like more. The order of the scenes (e.g., with or without emotion) was generated beforehand. The total number of volunteers was 256: 128 males, 126 females, and 2 that chose not to specify their gender. The average age was 27.29 years, with standard deviation of 16.58, minimum age was 4 and maximum 76. We accepted this variability as representative of a general population.

>>>>

 The case study presented in this paper was held at an exhibition with two main objectives:
 (i) Cross-validate the findings obtained from a previous experiment~\cite{Angel2017-2}, which was done to study the attribution of different linear and angular velocities, oscillation angle, direction, and orientation of the platform to express: \textit{Anger}, \textit{Happiness}, \textit{Sadness} and \textit{Fear}. A total of 196 value combinations were designed, 20 of which, randomly selected, were presented to each participant. Subjects had to select for each presentation the most representative term describing it among the four emotions and two mental states (i.e., \textit{Excitement} and \textit{Tenderness}), added as confounding terms. (ii) Verify whether participants would prefer scenes where the robot moves expressing emotions, or rather moves on the same trajectory without any emotion expression. To control scene variability two cameras and eight AR tags were used together with a Kalman filter to improve robot localization on the stage. AR tags were detected using the ROS package ar\_track\_alvar~\cite{artag2015}. The distribution of the cameras and the tags is reported in Figure~\ref{fig:setup_fourth}.

@@

@@

This case study was done at an exhibition during a period of two days. People coming to the stand were asked to volunteer to participate in this study, which was divided in to parts. In the first one, each subject was exposed to two emotion expressions. %TODO Was it so? I have turned the sentence... 
The two emotions were generated to guarantee a random uniform presentation. In the second part, participants were explained that a small scene was going to be presented twice, and they had to select the one they like more. The order of the scenes (e.g., with or without emotion) was generated beforehand. The total number of volunteers was 256: 128 males, 126 females, and 2 that chose not to specify their gender. The average age was 27.29 years, with standard deviation of 16.58, minimum age was 4 and maximum 76. We accepted this variability as representative of a general population.

### 2017-03-26 RO-MAN-2017-1/Parts/Introduction.tex
two ro-man articles compiled, ro-man2 still requires more work. let's see what the reviewers say about both articles

Emotions and mental states are not just expressed through facial expression and voice, but also by body postures and other features~\cite{Gelder2008}. Many psychological studies have focused their attention on how emotions are conveyed through the face~\cite{Ekman2004},~\cite{kleinsmith2012affective}. Similarly, the robotics community uses anthropomorphic faces (e.g.,~\cite{Arras2012},~\cite{Breazeal2002}) and bodies (e.g.,~\cite{Canamero2010},
%~\cite{haering2011},
~\cite{Destephe2013}
) to convey emotions.
In many situations, the presence of anthropomorphic elements would be out of place and not justified by the main robot's functionalities. Most of the current and future robotics platforms on the market will not require anthropomorphic faces~\cite{Breazeal2002} or limbs~\cite{Li2011}. For instance, floor cleaning robots are robots that do not have any anthropomorphic characteristics and still they accomplish their task.
This generates the necessity to study other mechanisms that could help to convey emotions, which could give people an idea about the robot's state, and engage the user in long term relations.

However,the amount of works studying non-anthropomorphic features in robotics (e.g.,~\cite{Saerbeck2010}
%,Lakatos2014,Sharma2013,Novika2015}
) remains still small in comparison to those exploiting anthropomorphic features. Moreover, these works do not provide specific range of values for the characteristics used to express the implemented emotions. For example Suk and collaborators~\cite{NAM2014} in their study give specific values for acceleration and curvature, but their connection to specific emotions is not described. Rather, they established a relationship between their values and pleasure/arousal dimensions. This could help to select specific features to convey certain emotions, but it would be better to know the precise range of values to be assigned to each characteristic to avoid misinterpretation with wrong emotions. These values could be used in social robotics to show emotions and, as consequence, increase their acceptance.

>>>>

Emotions and mental states are not just expressed through facial expression and voice, but also by body postures and other features~\cite{Gelder2008}. Many psychological studies have focused their attention on how emotions are conveyed through the face~\cite{kleinsmith2012affective}. Similarly, the robotics community uses anthropomorphic faces (e.g.~\cite{Breazeal2002}) and bodies (e.g.~\cite{Canamero2010}) to convey emotions.
In many situations, the presence of anthropomorphic elements would be out of place and not justified by the main robot's functionalities. Most of the current and future robotics platforms on the market will not require anthropomorphic faces~\cite{Breazeal2002} or limbs~\cite{Li2011}. For instance, floor cleaning robots are robots that do not have any anthropomorphic characteristics and still they accomplish their task.
This generates the necessity to study other mechanisms that could help to convey emotions, which could give people an idea about the robot's state, and engage the user in long term relations.

However,the amount of works studying non-anthropomorphic features in robotics (e.g.,~\cite{Saerbeck2010}) remains still small in comparison to those exploiting anthropomorphic features. Moreover, these works do not provide specific range of values for the characteristics used to express the implemented emotions. For example Suk and collaborators~\cite{NAM2014} in their study give specific values for acceleration and curvature, but their connection to specific emotions is not described. Rather, they established a relationship between their values and pleasure/arousal dimensions. This could help to select specific features to convey certain emotions, but it would be better to know the precise range of values to be assigned to each characteristic to avoid misinterpretation with wrong emotions. These values could be used in social robotics to show emotions and, as consequence, increase their acceptance.

### 2017-03-26 RO-MAN-2017-1/Parts/RelatedWork.tex
two ro-man articles compiled, ro-man2 still requires more work. let's see what the reviewers say about both articles

The first path is characterized by the use of cue positions to project emotions~\cite{NAO2013}. In some cases, special attention has been taken to determine head's angle and arms' position contribution to specific emotion expressions~\cite{Brown2014}. 
Nevertheless, current humanoid platforms cannot generate smooth gaits, and this limits the quality of body movements. To overcome this limitation, some researchers have reduced the human resemblance (e.g., eliminating limbs and facial expressions) to increase platform mobility and study new mechanisms to project emotions~\cite{Arras2012}. Reducing even more the anthropomorphism, Saerbeck and Christoph used a Roomba platform to study the contribution of curvature in a trajectory to conceive emotional states~\cite{Saerbeck2010}. Similarly Lourens and Barakova~\cite{BarakovaL10} implemented a set of behaviors to determine the emotion  perceived from diverse movements, which were selected from the work done by Camurri et al.~\cite{pop00002}. Continuing with their research on how robotics' behaviours are interpreted by people, Barakova and collaborators~\cite{Barakova2013} created a closet in which lights could be manipulated to convey pre-defined behaviours. The robot's behaviours were defined using the Interpersonal Behaviour Circle (ICB)~\cite{Leary57}, which is based on two dimensions (dominance-submission and hate-love). Their findings suggest that electronic systems can elicit a type of reactions different from the one expected by theories of interpersonal communication. Despite the type of platform used, these works offers an overview of features that could be used to express certain emotions, but they not give precise feature values to implement the emotions.

>>>>

The first path is characterized by the use of cue positions to project emotions~\cite{NAO2013}. In some cases, special attention has been taken to determine head's angle and arms' position contribution to specific emotion expressions~\cite{Brown2014}. 
Nevertheless, current humanoid platforms cannot generate smooth gaits, and this limits the quality of body movements. To overcome this limitation, some researchers have reduced the human resemblance (e.g., eliminating limbs and facial expressions) to increase platform mobility and study new mechanisms to project emotions~\cite{Arras2012}. Reducing even more the anthropomorphism, Saerbeck and Christoph used a Roomba platform to study the contribution of curvature in a trajectory to conceive emotional states~\cite{Saerbeck2010}. Similarly Lourens and Barakova~\cite{BarakovaL10} implemented a set of behaviors to determine the emotion  perceived from diverse movements, which were selected from the work done by Camurri et al.~\cite{pop00002}. Similarly, Barakova and collaborators~\cite{Barakova2013} created a closet in which lights could be manipulated to convey pre-defined behaviours. The robot's behaviours were defined using the Interpersonal Behaviour Circle (ICB). Their findings suggest that electronic systems can elicit a type of reactions different from the one expected by theories of interpersonal communication. Despite the type of platform used, these works offers an overview of features that could be used to express certain emotions, but they not give precise feature values to implement the emotions.

### 2017-03-26 RO-MAN-2017-1/RO-MAN2017.tex
two ro-man articles compiled, ro-man2 still requires more work. let's see what the reviewers say about both articles

% The following packages can be found on http:\\www.ctan.org
\usepackage{graphics} % for pdf, bitmapped graphics files
\usepackage{epsfig} % for postscript graphics files
\usepackage{subfigure}
\usepackage{multirow} 
\usepackage{url}
\usepackage{slashbox}
\newcommand{\co}{\cellcolor{gray!40}}
\usepackage[table]{xcolor}
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed
\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command

@@

\author{Julian M. Angel-Fernandez$^{1}$ and Andrea Bonarini$^{2}$% <-this % stops a space
\thanks{$^{1}$Julian M. Angel-Fernandez is a Research at Automation and Control Institute, Vienna University of Technology, Vienna, Austria
        {\tt\small julian.angel.fernandez@tuwien.ac.at}}%
\thanks{$^{2}$Andrea Bonarini is a Professor at the Department of Electronics, Information, and Bioengineering, Politecnico di Milano, Milan, Italy,
        {\tt\small andrea.bonarini@polimi.it}}%
}

>>>>

% The following packages can be found on http:\\www.ctan.org
\usepackage{graphics} % for pdf, bitmapped graphics files
\usepackage{epsfig} % for postscript graphics files
\usepackage{subfigure}
\usepackage{multirow} 
\usepackage{url}
\usepackage{slashbox}

\usepackage[table]{xcolor}
\newcommand{\co}{\cellcolor{gray!40}}
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed
\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command

@@

\author{Julian M. Angel-Fernandez$^{1}$ and Andrea Bonarini$^{2}$% <-this % stops a space
\thanks{$^{1}$Julian M. Angel-Fernandez is a Researcher at Automation and Control Institute, Vienna University of Technology, Vienna, Austria
        {\tt\small julian.angel.fernandez@tuwien.ac.at}}%
\thanks{$^{2}$Andrea Bonarini is a Professor at the Department of Electronics, Information, and Bioengineering, Politecnico di Milano, Milan, Italy,
        {\tt\small andrea.bonarini@polimi.it}}%
}

### 2017-03-26 RO-MAN-2017-2/Parts/EmotionalTree.tex
fiixed some grammar issues
two ro-man articles compiled, ro-man2 still requires more work. let's see what the reviewers say about both articles

Emotional Execution Tree is a connected acyclic graph $G(V,E)$ with $|V|$ vertexes and $|E|$ edges. The root and non-leaf nodes could be of either \textit{parallel} or \textit{sequential} type, which represent the type of actions that would be executed by the system. Sequential nodes execute one branch after the other. Once all the branches have been executed, asequential node notifies its predecessor to continue with the execution of other branches. The parallel node executes all the branches at the same time. This type of node offers two levels of synchronization: action and emotion. The action level synchronization enables the possibility to stop all the branches once the main branch has completed its execution. This main branch is established in the action message. On the other hand, the emotion level synchronization updates to a new emotional parameter of all the branches once the main branch updates to a new emotional parameter. %TODO The previous sentence is not clear. I cannot understando so I cannot fix it. What does it mean that updates to a parameter?
In both cases, any update or finish message will be ignored if the branch is not principal. These two levels of synchronization create four sub-types of parallel nodes: (i) action and emotion synchronous, (ii) action synchronous and emotion asynchronous, (iii) action asynchronous and emotion synchronous, or (iv) action and emotion asynchronous.

>>>>

Emotional Execution Tree is a connected acyclic graph $G(V,E)$ with $|V|$ vertexes and $|E|$ edges. The root and non-leaf nodes could be of either \textit{parallel} or \textit{sequential} type, which represent the type of actions that would be executed by the system. Sequential nodes execute one branch after the other. Once all the branches have been executed, a sequential node notifies its predecessor to continue with the execution of other branches. The parallel node executes all the branches at the same time. This type of node offers two levels of synchronization: action and emotion. The action level synchronization enables the possibility to stop all the branches once the main branch has completed its execution. This main branch is established in the action message. On the other hand, the emotion level synchronization makes all branches to shift to the next descriptor of the emotion. 
In both cases, any update or finish message will be ignored if the branch is not principal. These two levels of synchronization create four sub-types of parallel nodes: (i) action and emotion synchronous, (ii) action synchronous and emotion asynchronous, (iii) action asynchronous and emotion synchronous, or (iv) action and emotion asynchronous.

### 2017-03-26 RO-MAN-2017-2/Parts/Implementation.tex
two ro-man articles compiled, ro-man2 still requires more work. let's see what the reviewers say about both articles

%TODO in this section is missing a review on the terms and explain some concepts that are depicted in the figures.

The system was implemented in \textit{C++} and interfaced with ROS. The design (Fig.~\ref{fig:system_architecture}) was created following the description presented in the previous sections. 
\begin{figure*}
	\centering
	\includegraphics[width=1.0\textwidth]{Images/SystemArchitecture.png} 	
	\caption{General system design. Each simple action corresponds to one ROS node, and there is just one node for the emotion enrichment system. The ovals represent the ROS topic parameters, rectangles represent \textit{black boxes}, and texts outside containers represent input files that contain the system parametrization.}
	\label{fig:system_architecture}
\end{figure*}
The emotion enrichment core is divided in three different modules. 
Each module is responsible for one of the following phases:
\begin{enumerate}
	\item \textit{Generation of emotional execution tree:} this phase starts every time that a new action message is received. The process begins by parsing the format, verifying that the actions described on it exist in the system, and that the parameters correspond to the ones expected by each action included in the message. This parameters' verification is done on the implemented description for each action, which describes the parameters that are mandatory and those that are optional. %TODO Why some are mandatory and some not? What does this mean?
When the verification is done, and all the actions exist and the parameters correspond, an Emotional Execution Tree (EXT) such as the one presented in Figure~\ref{fig:emotional_enrichment} is created
	\item \textit{Emotion addition:} uses the EXT created in the previous phase. In this phase new simple actions ($sa$)
can be added to the EXT, and the $sa$'s parameters are modified following the emotion description, which is loaded from files. This process is broken down in two steps. First, all the actions that are required to convey the desired emotion, and that are not yet present are added. Second, the emotional parameters are modulated basing on the emotion's intensity and character traits. The final EXT is presented in Figure~\ref{fig:reference}.
	\begin{figure}
		\centering
		%Representation
	\includegraphics[width=0.45\textwidth]{./Images/exampleTreeE.png}
	\caption{Enrich Emotional Execution Tree.} 
	\label{fig:reference}
	\end{figure}
	\item \textit{Execution:} this is the last phase and it is done after the $ext$ is ''coloured'' with emotional characteristics (actions additions and emotional parameters). The decision to have two different communication channels, one for action parameters and another for the action emotional parameters, was taken to enable the possibility to update the emotional parameters without interfering with the current execution. In this phase a reference to the principal and emotional action is maintained, thus, when a new emotion is to be expressed, the system terminates the previous emotional action and starts executing the new ones without affecting the general execution of the actions. %TODO NOt clear. t seems that actually some action execution changes. What would you like to say?
\end{enumerate}

All the text message broadcast among the nodes are written using JSON format, which is human understandable, light, and can be parsed by many systems.\\ 
To test the system, two different platforms were used: Keepon Pro %TODO da citare la prima volta il paper descrittivo
and an own made platform called Triskarino.
Keepon has been used to test the interoperability of the system to different platforms, while Triskarino has been used in diverse case studies to explore emotion projection with a non anthropomorphic platform.
%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Keepon Test}
To test the system with Keepon (Fig.~\ref{fig:keepon}) was just necessary to implement the platform's controllers in each one of the simple action ROS nodes. Given that this platform does not have the capability to displace its body, the action move body was not implemented. Once added these controllers to the system, we proceeded to modify the configuration files to use this platform instead of Triskarino. With this small modification, we were able to change from one platform to other. In the video\footnote{YouTube video name Emotional Enrichment System, url: https://youtu.be/bRSXQ0rzkO8}
is shown the test where Keepon had to move its torso forward and backward to express a ``happy'' emotion. 
\begin{figure}
	\centering
	\includegraphics[width=0.2\textwidth]{./Images/Keepon.jpg}
	\caption{Keepon platform.}
	\label{fig:keepon}
\end{figure} 
The action is given by console telling the robot to bend the torso to a desired angle in x. The torso oscillation in y is added automatically by the system following the description given to happiness. 
 %%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Triskarino Test}
The system has been widely used with Triskarino (Fig.~\ref{fig:robot}) during our studies on projecting emotions with a non human like platform. 
\begin{figure}
	\centering
	\begin{subfigure}[c]{0.2\textwidth}
	\includegraphics[width=\textwidth]{./Images/upperFourthD.png}
	\caption{Triskarino design.}
	\label{fig:design}
	\end{subfigure}
	\begin{subfigure}[c]{0.2\textwidth}
	\includegraphics[width=\textwidth]{./Images/platform_fome.jpg}
	\caption{Triskarino with its cover.}
	\label{fig:triskar}
	\end{subfigure}
	\caption{Triskarino platform.}
	\label{fig:robot}
\end{figure}  
 During these studies the system was used in different experiments to move the robot on a straight line showing different emotions, each one selected from our command console. Moreover, to show the whole capabilities of the emotional enrichment system, a little scene was set up. The scene is a modification of the first part of the balcony scene of Shakespeare's Romeo and Juliet play~\cite{RAndJ}. The simplified sketch of our version could be seen in the Fig.~\ref{fig:triskar_test}. As it can be seen the stage was divided in 81 squares and the positions were given to the system in terms of the desire square to be reached, which allows the robot to adjust to the stage's dimension. %TODO I cannot understand what follows. What is a context?
The whole sequence of actions were specified as unique action, having several sequential context and just move body action. The other actions, such as oscillate body or blend upper body were added online accordingly the desired emotion.

>>>>

The system was implemented in \textit{C++} and interfaced with ROS. The design (Fig.~\ref{fig:system_architecture}) was created following the description presented in the previous sections. 
\begin{figure*}
	\centering
	\includegraphics[width=1.0\textwidth]{Images/SystemArchitecture.png} 	
	\caption{General system design. Each simple action corresponds to one ROS node, and there is just one node for the emotion enrichment system. The ovals represent the ROS topic parameters, rectangles represent \textit{black boxes}, and texts outside containers represent input files that contain the system parametrization.}
	\label{fig:system_architecture}
\end{figure*}
The emotion enrichment core is divided in three different modules. 
Each module is responsible for one of the following phases:
\begin{enumerate}
	\item \textit{Generation of emotional execution tree:} this phase starts every time that a new action message is received. The process begins by parsing the format, verifying that the actions described on it exist in the system, and that the parameters correspond to the ones expected by each action included in the message. 
When the verification is done, and all the actions exist and the parameters correspond, an EXT such as the one presented in Figure~\ref{fig:emotional_enrichment} is created.
	\item \textit{Emotion addition:} uses the EXT created in the previous phase. In this phase new simple actions ($sa$)
can be added to the EXT, and the $sa$'s parameters are modified following the emotion description, which is loaded from files. This process is broken down in two steps. First, all the actions that are required to convey the desired emotion, and that are not yet present are added. Second, the emotional parameters are modulated basing on the emotion's intensity and character traits. The final EXT is presented in Figure~\ref{fig:reference}.
	\begin{figure}
		\centering
		%Representation
	\includegraphics[width=0.45\textwidth]{./Images/exampleTreeE.png}
	\caption{EXT enriched with emotion for the EXT presented in Figure~\ref{fig:emotional_enrichment}.} 
	\label{fig:reference}
	\end{figure}
	\item \textit{Execution:} this is the last phase and it is done after the EXT is ''coloured'' with emotional characteristics (actions additions and emotional parameters). The decision to have two different communication channels, one for action parameters and another for the action emotional parameters, was taken to enable the possibility to update the emotional parameters without interfering with the current execution. 
\end{enumerate}

All the text message broadcast among the nodes are written using JSON format, which is human understandable, light, and can be parsed by many systems.\\ 
To test the system, two different platforms were used: Keepon Pro %TODO da citare la prima volta il paper descrittivo
and an own made platform called Triskarino.
Keepon has been used to test the interoperability of the system to different platforms, while Triskarino has been used in diverse case studies to explore emotion projection with a non anthropomorphic platform.
%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Keepon Test}
To test the system with Keepon (Fig.~\ref{fig:keepon}) was just necessary to implement the platform's controllers in each one of the simple action ROS nodes. Given that this platform does not have the capability to displace its body, the action move body was not implemented. Once added these controllers to the system, we proceeded to modify the configuration files to use this platform instead of Triskarino. With this small modification, we were able to change from one platform to other. In the video\footnote{YouTube video name Emotional Enrichment System, url: https://youtu.be/bRSXQ0rzkO8}
is shown the test where Keepon had to move its torso forward and backward to express a ``happy'' emotion. 
\begin{figure}
	\centering
	\includegraphics[width=0.2\textwidth]{./Images/Keepon.jpg}
	\caption{Keepon platform.}
	\label{fig:keepon}
\end{figure} 
The action is given by console telling the robot to bend the torso to a desired angle in x. The torso oscillation in y is added automatically by the system following the description given to happiness. 
 %%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Triskarino Test}
The system has been widely used with Triskarino (Fig.~\ref{fig:robot}) during our studies on projecting emotions with a non human like platform. 
\begin{figure}
	\centering
	\begin{subfigure}[c]{0.2\textwidth}
	\includegraphics[width=\textwidth]{./Images/upperFourthD.png}
	\caption{Triskarino design.}
	\label{fig:design}
	\end{subfigure}
	\begin{subfigure}[c]{0.2\textwidth}
	\includegraphics[width=\textwidth]{./Images/platform_fome.jpg}
	\caption{Triskarino with its cover.}
	\label{fig:triskar}
	\end{subfigure}
	\caption{Triskarino platform.}
	\label{fig:robot}
\end{figure}  
 During these studies the system was used in different experiments to move the robot on a straight line showing different emotions, each one selected from our command console. Moreover, to show the whole capabilities of the emotional enrichment system, a little scene was set up. The scene is a modification of the first part of the balcony scene of Shakespeare's Romeo and Juliet play~\cite{RAndJ}. The simplified sketch of our version could be seen in the Fig.~\ref{fig:triskar_test}. As it can be seen the stage was divided in 81 squares and the positions were given to the system in terms of the desire square to be reached, which allows the robot to adjust to the stage's dimension.
The whole sequence of actions were specified as unique action, having several sequential nodes and just move body action. The other actions, such as oscillate body or blend upper body were added online accordingly the desired emotion.

### 2017-03-26 RO-MAN-2017-2/Parts/SimpleActions.tex
two ro-man articles compiled, ro-man2 still requires more work. let's see what the reviewers say about both articles

In order to generate a system that could be used in diverse platforms, an abstraction level in which all of them could fall is required. Therefore simple and compound actions are used to achieve this goal. Simple actions are actions that are considered as primitives: they are used as building blocks. As a consequence, these actions are described in the system and are the ones whose parameters are changed to embed the emotions. Their description specifies mandatory and optional parameters that are required to execute an action. These parameters are modified accordingly to the information provided by the emotional parameters. Compound actions are actions that are created from simple actions and other compound actions. These actions are not implemented in the system, but, if it is needed, they can be described in it (e.g., compound actions that are used often). %TODO it is not clear what does it mean that they are not implemented.

>>>>

In order to generate a system that could be used in diverse platforms, an abstraction level in which all of them could fall is required. Therefore simple and compound actions are used to achieve this goal. Simple actions are actions that are considered as primitives: they are used as building blocks. As a consequence, these actions are described in the system and are the ones whose parameters are changed to embed the emotions. Their description specifies mandatory and optional parameters that are required to execute an action. These parameters are modified accordingly to the information provided by the emotional parameters. Compound actions are actions that are created from simple actions and other compound actions. For example, two actions that are executed in parallel are considered as compound actions. These actions are not always implemented in the system. They can be described in the action messages without being implemented. In case that a compound action is used often, it could be implemented in the system. The implementation describes action's parameters, other compound actions, simple actions and mapping between actions and parameters.

### 2017-03-26 RO-MAN-2017-2/Parts/description.tex
two ro-man articles compiled, ro-man2 still requires more work. let's see what the reviewers say about both articles

The description files allow the parametrization and extent the system to fit other applications is required. To standarize the mechanisms to describe these files, it was decided to create a language to describe them. This allows other researchers to create their own files and to extend the language to fit their needs. It was decided to use the Extended Backus-NAur Form (EBN)F to describe this language.
\subsection{Emotion Description} 
The language created to describe emotion descriptors is the following:
\begin{small}

@@

\endgrammar
\end{small}
%TODO give information regardin 1) why repetition 2) why a set of parameters 3) why use of time and space

@@

%TODO give information regarding: primary and emotion synch

>>>>

The description files allow the parametrization and extent the system to fit other applications is required. To standarize the mechanisms to describe these files, it was decided to create a language to describe them. This allows other researchers to create their own files and to extend the language to fit their needs. It was decided to use the Extended Backus-Naur Form (EBNF) to describe this language.
\subsection{Emotion Description} 
The language created to describe emotion descriptors is the following:
\begin{small}

@@

\endgrammar
\end{small}
The attribute repetition is used to specify if the descriptors of an emotion parameters must be executed in an infinite loop. As it was already mentioned, the emotional parameters implemented for this version correspond to the move actions.

@@

### 2017-03-26 RO-MAN-2017-2/RO-MAN2017-2.tex
two ro-man articles compiled, ro-man2 still requires more work. let's see what the reviewers say about both articles

\subsection{Emotion Parameters}
The emotion parameters specify how actions parameters should be modified in order to express an emotion. However, there is going to be a diverse set of parameters depending on the nature of the action. For example, consider the actions \textit{move body} and \textit{speak}. The human action of speaking is related to changes in the vocal cords to produce different sounds, while \textit{move body} is directly related to the body movement. The speaking action would have the following parameters: text, pitch and tone. On the other hand, the action \textit{move body} would have as parameters the desired destination, velocity, and trajectory constraints. Since each action is characterized by different parameters, their modification must be done accordingly to the parameters. Therefore, a variety of emotional parameters would be necessary to describe the modification of all type of actions. So, it was decided to focus on the parameters that could be used to express emotions with the robots we have considered: Triskarino and Keepon. It was determined that these parameters could just use movement actions, which are described by position, velocity and acceleration.    
\subsection{Simple and Compound Actions}
\input{./Parts/SimpleActions} 
\section{Emotional Tree}
\label{sec:emotional_execution_tree}
\input{./Parts/EmotionalTree}
\section{Describing Components}
\label{sec:description}
\input{./Parts/description}
\section{Implementation}
\label{sec:implementation}
\input{./Parts/Implementation}
\section{Conclusions and Further work}
An Emotional Enrichment System has been designed and implemented to enrich robots' movements with emotions. To achieve this,  an Emotional Execution Tree created from three different types of nodes was used: simple actions, parallel, and sequential. Simple actions are functions that map a set of parameters to specific movements. Sequential nodes execute the sequence of actions associated to this type of node in a sequential order, while parallel executes them all at the same time. To enable synchronization among simple actions, parallel nodes could be one of four different subtypes, and sequential just one of two different subtypes. A formalization of the Emotional Execution Tree and the principal considerations during the implementation of the system have also been provided. To show the system's versatility, it was used with quite different platforms such as a Keepon Pro and Triskarino. Keepon was used to perform simple actions, while Triskarino was used to test complex actions with different parametrizations of emotions.

>>>>

\subsection{Emotion Parameters}
The emotion parameters specify how actions parameters should be modified in order to express an emotion. These parameters could include a sequence of descriptors. For example, these descriptors could be used to generate changes of velocity over time or position to express an emotion. However, there is going to be a diverse set of parameters and descriptors depending on the nature of the action. For example, consider the actions \textit{move body} and \textit{speak}. The human action of speaking is related to changes in the vocal cords to produce different sounds, while \textit{move body} is directly related to the body movement. The speaking action would have the following parameters: text, pitch and tone. On the other hand, the action \textit{move body} would have as parameters the desired destination, velocity, and trajectory constraints. However, in order to express emotion in speak action, it is required to modify pitch variables (i.e. fundamental frequency level, contour and jitter), intensity and speech rate~\cite{Gunes2011}. On the other hand, changing linear velocity and adding oscillation could be used to convey emotion in the move body action~\cite{Angel2016c}. Since each action requires modification on diverse parameters, a variety of emotional parameters would be necessary to describe the modification of all type of actions. So, it was decided to focus on the parameters that could be used to express emotions with the robots we have considered: Triskarino and Keepon. It was determined that these parameters could just use movement actions, which are described by position, velocity and acceleration.    
\subsection{Simple and Compound Actions}
\input{./Parts/SimpleActions} 
\section{Emotional Tree}
\label{sec:emotional_execution_tree}
\input{./Parts/EmotionalTree}
\section{Describing Components}
\label{sec:description}
\input{./Parts/description}
\section{Implementation}
\label{sec:implementation}
\input{./Parts/Implementation}
\section{Conclusions and Further work}
An Emotional Enrichment System has been designed and implemented to enrich robots' movements with emotions. To achieve this,  an Emotional Execution Tree created from three different types of nodes was used: simple actions, parallel, and sequential. Simple actions are functions that map a set of parameters to specific movements. Sequential nodes execute the sequence of actions associated to this type of node in a sequential order, while parallel executes them all at the same time. To enable synchronization among simple actions, parallel nodes could be one of four different subtypes, and sequential just one of two different subtypes. A formalization of the Emotional Execution Tree and the principal considerations during the implementation of the system have also been provided. To show the system's versatility, it was used with quite different platforms such as a Keepon Pro and Triskarino. Keepon was used to perform simple actions, while Triskarino was used to test complex actions with different parametrizations of emotions.
